---
title: "Assignment3"
author: "RF"
date: "3/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Multivariate linear models

In this assignment you will investigate the relation between different aspects of IQ and symptom severity in ASD (as measured by ADOS). The assignment is meant to make you practice on linear models, multiple predictors and interactions.

The data you will use is the outcome of the psychological testing of the children you analyzed in methods 3.
Data: https://www.dropbox.com/s/hom2qnmp9hyfwmu/Assignment3Data.csv?dl=0
The data consists of ChildID, gender, age, diagnosis (ASD), symptom severity (ADOS), several aspects of IQ (NonVerbal/Perceptual,Verbal,Social) and the visit at which the tests were administered. The questions should be answered by focusing on visit 1.

The questions you will have to answer are:

1. Assess the relation between symptom severity and IQ (focus on visit 1 and children with ASD) and report the model, a plot of the model, a couple of lines describing the quality of the model and interpreting the results. P.S. Should you scale? P.P.S. Don't forget to motivate your priors. P.P.P.S. At least one plot for results and a plot for quality of each model (here and in the next questions) would be appreciated.
1.1. Verbal IQ and ADOS
1.2. Non Verbal IQ and ADOS
1.3. Social IQ and ADOS

```{r}
#Load libraries
library(dplyr)
library(rethinking)
library(ggplot2)

#Load data
d_full = read.csv("Assignment3Data.csv")
#Set to only Visit 1
d_full = filter (d_full, Visit==1)
#Remove NA's
d_full = d_full[complete.cases(d_full),]

#Center all relevant variables
d_full$ADOS.s = scale(d_full$ADOS, scale = F, center = T)[,]
d_full$VerbalIQ.s = scale(d_full$VerbalIQ, scale = F, center = T)[,]
d_full$NonVerbalIQ.s = scale(d_full$NonVerbalIQ, scale = F, center = T)[,]
d_full$SocialIQ.s = scale(d_full$SocialIQ, scale = F, center = T)[,]

#Filter for visit 1, only ASD
d = subset(d_full, ASD == 1)

```

```{r}
#FOR VERBAL IQ
#Create model
m = map(
  alist(
    ADOS ~ dnorm (mu, sigma),
    mu <- a + b*VerbalIQ ,
    a ~ dnorm (15, 15),
    b ~ dnorm (0, 5),
    sigma ~ dunif (0,10)
  ) ,
  data = d
)

#Get results of model
precis (m, corr=T)

#Plot results
#Create many different posteriors
post <- extract.samples(m) 

#a function that calculates mu for a specific x-value
mu.link <- function(x) post$a + post$b*x
#Make a list of numbers for VerbalIQ
Predictor_sequence <- seq( from=min(d$VerbalIQ) , to=max(d$VerbalIQ) , by=0.1 )
#Calculate the mu for each of the verbalIQ's
mu <- sapply( Predictor_sequence , mu.link )
#What's the mean mu?
mu_mean <- apply( mu , 2 , mean )
#High Probability Density Intervals
mu_HPDI <- apply( mu , 2 , HPDI , prob=0.89 )

#Simulate mean outputs
Output_sim <- sim( m , data=list(VerbalIQ=Predictor_sequence) )

#Create percentile interval - the 89% most likely area
Output_PI <- apply( Output_sim , 2 , PI , prob=0.89 )

#Plot the raw data
plot( ADOS ~ VerbalIQ , d , col=col.alpha(rangi2,0.5) )
# the MAP model line
lines( Predictor_sequence , mu_mean )
# draw high probability density intervals for line
shade( mu_HPDI , Predictor_sequence ) 
# draw PI region for simulated heights
shade( Output_PI , Predictor_sequence )


###ASSESSING QUALITY:
dens(Output_sim, col = "red", xlim = c(0, 30), ylim = c(0,0.12),  xlab = "ADOS")
par(new=TRUE)
dens(d$ADOS, xlim = c(0, 30), ylim = c(0,0.12), xlab = "ADOS")

dens(mu - d$ADOS)

```

```{r}
#FOR NON-VERBAL IQ
#Create model
m = map(
  alist(
    ADOS ~ dnorm (mu, sigma),
    mu <- a + b*NonVerbalIQ ,
    a ~ dnorm (15, 15),
    b ~ dnorm (0, 5),
    sigma ~ dunif (0,10)
  ) ,
  data = d
)

#Get results of model
precis (m, corr=T)

#Plot results
#Create posterior
post <- extract.samples(m) 

#a function that calculates mu fora specific x-value
mu.link <- function(x) post$a + post$b*x
#Make a list of numbers for VerbalIQ
Predictor_sequence <- seq( from=min(d$NonVerbalIQ) , to=max(d$NonVerbalIQ) , by=0.1 )
#Calculate the mu for each of the verbalIQ's
mu <- sapply( Predictor_sequence , mu.link )
#What's the men mu?
mu_mean <- apply( mu , 2 , mean )
#High Probability Density Intervals
mu_HPDI <- apply( mu , 2 , HPDI , prob=0.89 )

#Simulate mean outputs
Output_sim <- sim( m , data=list(NonVerbalIQ=Predictor_sequence) )

#Create probability intervals - between 89
Output_PI <- apply( Output_sim , 2 , PI , prob=0.89 )

#Plot the raw data
plot( ADOS ~ NonVerbalIQ , d , col=col.alpha(rangi2,0.5) )
# the MAP model line
lines( Predictor_sequence , mu_mean )
# draw high probability density intervals for line
shade( mu_HPDI , Predictor_sequence ) 
# draw PI region for simulated heights
shade( Output_PI , Predictor_sequence ) #89 PI

###ASSESSING QUALITY:
#Plot prediction density agains observed density
dens(Output_sim, col = "red", xlim = c(0, 30), ylim = c(0,0.12),  xlab = "ADOS")
par(new=TRUE)
dens(d$ADOS, xlim = c(0, 30), ylim = c(0,0.12), xlab = "ADOS")

#Plot residuals
dens(mu - d$ADOS)
```

```{r}
#FOR SOCIAL IQ
#Create model
m = map(
  alist(
    ADOS ~ dnorm (mu, sigma),
    mu <- a + b*SocialIQ ,
    a ~ dnorm (15, 15),
    b ~ dnorm (0, 5),
    sigma ~ dunif (0,10)
  ) ,
  data = d
)

#Get results of model
precis (m, corr=T)

###Create posterior
post <- extract.samples(m) 

###Crete full predictions
#Make a list of numbers for VerbalIQ
Predictor_sequence <- seq( from=min(d$SocialIQ) , to=max(d$SocialIQ) , by=0.1 )

#Make function that calculates mu for each specific predictor-value
mu.link <- function(x) post$a + post$b*x

#Calculate the mu for each of the predictors created steps
mu <- sapply( Predictor_sequence , mu.link )
#What's the mean mu?
mu_mean <- apply( mu , 2 , mean )
#High Probability Density Intervals
mu_HPDI <- apply( mu , 2 , HPDI , prob=0.89 )

#
#Simulate full outputs
Output_sim <- sim( m , data=list(SocialIQ=Predictor_sequence) )

#Create probability intervals - between 89
Output_PI <- apply( Output_sim , 2 , PI , prob=0.89 )

#Plot the raw data
plot( ADOS ~ SocialIQ , d , col=col.alpha(rangi2,0.5) )
# the MAP model line
lines( Predictor_sequence , mu_mean )
# The high probability intervals
shade( mu_HPDI , Predictor_sequence ) 
# And the PI region for simulated heights
shade( Output_PI , Predictor_sequence ) #89 PI

###ASSESSING QUALITY:
dens(Output_sim, col = "red", xlim = c(0, 30), ylim = c(0,0.12),  xlab = "ADOS")
par(new=TRUE)
dens(d$ADOS, xlim = c(0, 30), ylim = c(0,0.12), xlab = "ADOS")

#Plot residuals
dens(mu - d$ADOS)


```



2. Do the different aspects of IQ account for different portions of the variance in ADOS? 
2.1. Does it make sense to have all IQ measures in the same model? First write a few lines answering the question and motivating your answer, including a discussion as to what happens when you put all of them in the same model. Then build a model following your answer. If your answer is "no", you are not free, you still have to answer: are there alternative ways of answering the question?
2.2. Build the model, assess its quality, write a few lines interpreting the results.


```{r}
#All IQ's
#Create model
m = map(
  alist(
    ADOS ~ dnorm (mu, sigma),
    mu <- a + b1*VerbalIQ.s + b2*NonVerbalIQ.s + b3*SocialIQ.s,
    a ~ dnorm (0, 15),
    b1 ~ dnorm (0, 5),
    b2 ~ dnorm (0, 5),
    b3 ~ dnorm (0, 5),
    sigma ~ dunif (0,10)
  ) ,
  data = d
)

#Get results of model
precis (m, corr=T)
plot(precis (m, corr=T))

#Predictive Posterior check
Output_sim <- sim( m, data=d)

#Plotting predicted and observed distributions
dens(Output_sim, col = "red", xlim = c(0, 30), ylim = c(0,0.12),  xlab = "ADOS")
par(new=TRUE)
dens(d$ADOS, xlim = c(0, 30), ylim = c(0,0.12), xlab = "ADOS")

#Plotting residual
dens(Output_sim - d$ADOS)

```

3. Let's now include also the TD children. Does it make sense to ask whether IQ and ADOS are related? Motivate your answer. In any case, if you wanted to build a model to answer that question, which model would you build? Run the model, assess its quality, write a few lines interpreting the results.


```{r}

#All IQ's
#Create model
m = map(
  alist(
    ADOS ~ dnorm (mu, sigma),
    mu <- a + bv*VerbalIQ.s + bnv*NonVerbalIQ.s + bs*SocialIQ.s,
    a ~ dnorm (0, 15),
    bv ~ dnorm (0, 5),
    bnv ~ dnorm (0, 5),
    bs ~ dnorm (0, 5),
    sigma ~ dunif (0,10)
  ) ,
  data = d_full
)

#Get results of model
precis (m, corr=T)
plot(precis (m, corr=T))


#Predictive posterior check
Output_sim <- sim( m, data=d_full)

dens(Output_sim, col = "red", xlim = c(-20, 30), ylim = c(0,0.13),  xlab = "ADOS")
par(new=TRUE)
dens(d_full$ADOS, xlim = c(-20, 30), ylim = c(0,0.13), xlab = "ADOS")

```

```{r}
#With Diagnosis
#Create model
m = map(
  alist(
    ADOS ~ dnorm (mu, sigma),
    mu <- a + bv*VerbalIQ.s + bnv*NonVerbalIQ.s + bs*SocialIQ.s + bD*ASD
    + bDv*VerbalIQ.s*ASD + bDnv*NonVerbalIQ.s*ASD + bDs*SocialIQ.s*ASD,
    a ~ dnorm (0, 15),
    bv ~ dnorm (0, 5),
    bnv ~ dnorm (0, 5),
    bs ~ dnorm (0, 5),
    bD ~ dnorm (0, 5),
    bDv ~ dnorm (0, 5),
    bDnv ~ dnorm (0, 5),
    bDs ~ dnorm (0, 5),
    sigma ~ dunif (0,10)
  ) ,
  data = d_full
)

#Get results of model
precis (m, corr=T)
plot(precis (m, corr=T))

#Predictive posterior check
Output_sim <- sim( m, data=d_full)

dens(Output_sim, col = "red", xlim = c(-20, 30), ylim = c(0,0.13),  xlab = "ADOS")
par(new=TRUE)
dens(d_full$ADOS, xlim = c(-20, 30), ylim = c(0,0.13), xlab = "ADOS")

```



4. Let's discuss contents:
4.1. You have three scores for IQ, do they show shared variance? Is that the same in TD and ASD? What does that tell us about IQ?
4.2. You have explored the relation between IQ and ADOS. How do you explain that relation from a cognitive perspective? N.B. You can present alternative hypotheses.

5. Bonus questions: Including measurement errors. 
5.1. Let's assume that ADOS has a measurement error of 1. How would you model that in any of the previous models? 
5.2. We know that IQ has an estimated measurement error of 2.12. How would you include that? 


